\LARGE\bf \qustop: Quantum state optimizer

### 2021-05-04 docs/paper/main.tex
Adding more to paper.

\documentclass[nofootinbib,superscriptaddress,a4paper,twocolumn,longbibliography,floatfix,pra]{revtex4-2}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}

\usepackage{enumitem}

\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{physics}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage[left=23mm,right=13mm,top=35mm,columnsep=15pt]{geometry} 
\usepackage{adjustbox}
\usepackage{placeins}
\usepackage[T1]{fontenc}
\usepackage{lipsum}
\usepackage{csquotes}

\usepackage{listings}
\usepackage{xcolor}

@@

\usepackage{xspace}
\newcommand{\expect}[2]{\ensuremath{\langle #1 \rangle_{#2}}}

\setlength{\parindent}{0cm}

\usepackage[pdftex, pdftitle={Article}, pdfauthor={Author}]{hyperref} % For hyperlinks in the PDF
\usepackage{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}

@@

\newcommand{\tinyspace}{\mspace{1mu}}
\newcommand{\microspace}{\mspace{0.5mu}}
\newcommand{\negsmallspace}{\mspace{-1.5mu}}
\newcommand{\pt}{\operatorname{T}}
\newcommand{\ppt}{\operatorname{PPT}}
\newcommand{\sep}{\operatorname{Sep}}
\renewcommand{\int}{\operatorname{int}}
\renewcommand{\det}{\operatorname{Det}}
\renewcommand{\vec}{\operatorname{vec}}
\newcommand{\fid}{\operatorname{F}}
\newcommand{\im}{\operatorname{im}}

\renewcommand{\tr}{\operatorname{Tr}}
\renewcommand{\t}{{\scriptscriptstyle\mathsf{T}}}

\renewcommand{\abs}[1]{\lvert #1 \rvert}
\newcommand{\bigabs}[1]{\bigl\lvert #1 \bigr\rvert}
\newcommand{\Bigabs}[1]{\Bigl\lvert #1 \Bigr\rvert}
\newcommand{\biggabs}[1]{\biggl\lvert #1 \biggr\rvert}
\newcommand{\Biggabs}[1]{\Biggl\lvert #1 \Biggr\rvert}

\renewcommand{\ip}[2]{\langle #1 , #2\rangle}
\newcommand{\bigip}[2]{\bigl\langle #1, #2 \bigr\rangle}
\newcommand{\Bigip}[2]{\Bigl\langle #1, #2 \Bigr\rangle}
\newcommand{\biggip}[2]{\biggl\langle #1, #2 \biggr\rangle}
\newcommand{\Biggip}[2]{\Biggl\langle #1, #2 \Biggr\rangle}

@@

\newcommand{\bignorm}[1]{\bigl\lVert\tinyspace #1 \tinyspace\bigr\rVert}
\newcommand{\Bignorm}[1]{\Bigl\lVert\tinyspace #1 \tinyspace\Bigr\rVert}
\newcommand{\biggnorm}[1]{\biggl\lVert\tinyspace #1 \tinyspace\biggr\rVert}
\newcommand{\Biggnorm}[1]{\Biggl\lVert\tinyspace #1 \tinyspace\Biggr\rVert}

@@

@@

@@

\newcommand{\I}{\mathbb{I}}

@@

\newcommand{\complex}{\mathbb{C}}
\renewcommand{\natural}{\mathbb{N}}
\newcommand{\integer}{\mathbb{Z}}

\usepackage{FiraSans}
\newcommand{\poly}{\mathit{poly}}
\newcommand{\toqitofont}{%
	\fontfamily{FiraSans}%
	\selectfont}

\newcommand{\toqito}{ $|${\toqitofont toqito}$\rangle$\xspace}
\newcommand{\numpy}{\texttt{numpy}\xspace}
\newcommand{\tequila}{\textsc{toqito}}
\newenvironment{mylist}[1]{\begin{list}{}{
	\setlength{\leftmargin}{#1}
	\setlength{\rightmargin}{0mm}
	\setlength{\labelsep}{2mm}
	\setlength{\labelwidth}{8mm}
	\setlength{\itemsep}{0mm}}}
	{\end{list}}

@@

\newcommand{\yes}{\text{yes}}
\newcommand{\no}{\text{no}}

@@

\def\BB84{\mathsf{BB84}}
\def\CHSH{\mathsf{CHSH}}
\def\MUB{\mathsf{MUB}}
\def\qustop{\texttt{qustop}}

\begin{document}
\title{\qustop: Quantum state optimizer}

\author{Vincent Russo}
\email[E-mail:]{vincentrusso1@gmail.com}
\affiliation{-}

\date{\today}

\begin{abstract}
\end{abstract}

@@

%-----------------------------------------------------------------------------%
\section{Introduction}
\label{sec:introduction}
%-----------------------------------------------------------------------------%

Documentation and tutorials are available on the $\qustop$
\href{https://github.com/vprusso/toqito}{GitHub repository}~\cite{toqito}.

%-----------------------------------------------------------------------------%
\section{Fundamental objects}
\label{sec:fundemental_objects}
%-----------------------------------------------------------------------------%

%-----------------------------------------------------------------------------%
\subsection{Quantum states and ensembles}\label{sec:ensembles}
%-----------------------------------------------------------------------------%

%-----------------------------------------------------------------------------%
\section{Applications of qustop}
\label{sec:applications}
%-----------------------------------------------------------------------------%

%-----------------------------------------------------------------------------%
\subsection{Calculating optimal probabilities of distinguishing quantum states}\label{sec:opt_dist}
%-----------------------------------------------------------------------------%

%-----------------------------------------------------------------------------%
\subsection{Global distinguishability}\label{sec:global}
%-----------------------------------------------------------------------------%

%-----------------------------------------------------------------------------%
\subsection{PPT distinguishability}\label{sec:ppt}
%-----------------------------------------------------------------------------%

%-----------------------------------------------------------------------------%
\subsection{Separable distinguishability}\label{sec:separable}
%-----------------------------------------------------------------------------%

%-----------------------------------------------------------------------------%
\subsection{Calculating optimal probabilities of excluding quantum states}\label{sec:opt_exclude}
%-----------------------------------------------------------------------------%

%-----------------------------------------------------------------------------%
\subsection{Calculating optimal probabilities of cloning quantum states}\label{sec:opt_clone}
%-----------------------------------------------------------------------------%

\begin{figure}[!htpb]
    \centering
    \lstinputlisting[language=Python]{code/indstinguishable_mes.py}        
    \caption{Calculate the minimum-error probability of optimally
    distinguishing the states from Equation~\eqref{eq:ydy} via PPT
    measurements.}
    \label{fig:ppt_ydy}
\end{figure}

%-----------------------------------------------------------------------------%
\subsubsection{Quantum state discrimination via separable measurements}
\label{sec:}
%-----------------------------------------------------------------------------%

%-----------------------------------------------------------------------------%
\section{Conclusion}
%-----------------------------------------------------------------------------%

%-----------------------------------------------------------------------------%
\section{Acknowledgements}
%-----------------------------------------------------------------------------%

\bibliography{main}
\clearpage
\end{document}

>>>>

% pdflatex -s main.tex && bibtex main.aux && pdflatex -s main.tex -o main.pdf && open main.pdf
\documentclass[11pt]{article}

%-----------------------------------------------------------------------------%
% Fonts and symbols:
%-----------------------------------------------------------------------------%

\usepackage{mathpazo}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{mathtools}
\usepackage{listings}

%-----------------------------------------------------------------------------%
% Margins and page layout:
%-----------------------------------------------------------------------------%

\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{pdfpagemode=UseNone}
\frenchspacing

%-----------------------------------------------------------------------------%
% Theorem-like environments:
%-----------------------------------------------------------------------------%

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{observation}{Observation}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{prop}[theorem]{Proposition}
\newtheorem{cor}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}

%-----------------------------------------------------------------------------%
% Other packages:
%-----------------------------------------------------------------------------%

\usepackage{authblk}
\usepackage{tikz}
\usetikzlibrary{calc,positioning}
\usepackage{mdframed}

\mdfdefinestyle{figstyle}{
  linecolor=black!7,
  backgroundcolor=black!7,
  innertopmargin=10pt,
  innerleftmargin=25pt,
  innerrightmargin=25pt,
  innerbottommargin=10pt
}

\definecolor{White}{rgb}{1,1,1}
\definecolor{Black}{rgb}{0,0,0}
\definecolor{LightGray}{rgb}{.81,.81,.81}
\colorlet{ChannelColor}{LightGray}
\colorlet{ChannelTextColor}{Black}
\colorlet{ReadoutColor}{White}

\usepackage{xcolor}

@@

@@

\newcommand{\tinyspace}{\mspace{1mu}}
\newcommand{\microspace}{\mspace{0.5mu}}
\newcommand{\negsmallspace}{\mspace{-1.5mu}}
\newcommand{\op}[1]{\operatorname{#1}}
\newcommand{\tr}{\operatorname{Tr}}
\newcommand{\pt}{\operatorname{T}}
\newcommand{\POS}{\operatorname{POS}}
\newcommand{\PPT}{\operatorname{PPT}}
\newcommand{\SEP}{\operatorname{SEP}}
\newcommand{\rank}{\operatorname{rank}}
\renewcommand{\int}{\operatorname{int}}
\renewcommand{\det}{\operatorname{Det}}
\renewcommand{\vec}{\operatorname{vec}}
\newcommand{\fid}{\operatorname{F}}
\newcommand{\im}{\operatorname{im}}

\renewcommand{\t}{{\scriptscriptstyle\mathsf{T}}}

\newcommand{\abs}[1]{\lvert #1 \rvert}
\newcommand{\bigabs}[1]{\bigl\lvert #1 \bigr\rvert}
\newcommand{\Bigabs}[1]{\Bigl\lvert #1 \Bigr\rvert}
\newcommand{\biggabs}[1]{\biggl\lvert #1 \biggr\rvert}
\newcommand{\Biggabs}[1]{\Biggl\lvert #1 \Biggr\rvert}

\newcommand{\ip}[2]{\langle #1 , #2\rangle}
\newcommand{\bigip}[2]{\bigl\langle #1, #2 \bigr\rangle}
\newcommand{\Bigip}[2]{\Bigl\langle #1, #2 \Bigr\rangle}
\newcommand{\biggip}[2]{\biggl\langle #1, #2 \biggr\rangle}
\newcommand{\Biggip}[2]{\Biggl\langle #1, #2 \Biggr\rangle}

@@

\newcommand{\norm}[1]{\lVert\tinyspace #1 \tinyspace\rVert}
\newcommand{\bignorm}[1]{\bigl\lVert\tinyspace #1 \tinyspace\bigr\rVert}
\newcommand{\Bignorm}[1]{\Bigl\lVert\tinyspace #1 \tinyspace\Bigr\rVert}
\newcommand{\biggnorm}[1]{\biggl\lVert\tinyspace #1 \tinyspace\biggr\rVert}
\newcommand{\Biggnorm}[1]{\Biggl\lVert\tinyspace #1 \tinyspace\Biggr\rVert}

@@

\newcommand{\ket}[1]{
  \lvert\microspace #1 \microspace \rangle}

@@

\newcommand{\bra}[1]{
  \langle\microspace #1 \microspace \rvert}

@@

\newcommand{\I}{\mathbb{1}}

@@

\newcommand{\complex}{\mathbb{C}}
\newcommand{\real}{\mathbb{R}}
\renewcommand{\natural}{\mathbb{N}}
\newcommand{\integer}{\mathbb{Z}}

\newcommand{\qustop}{\texttt{qustop}}

\newcommand{\poly}{\mathit{poly}}

\newenvironment{mylist}[1]{\begin{list}{}{
	\setlength{\leftmargin}{#1}
	\setlength{\rightmargin}{0mm}
	\setlength{\labelsep}{2mm}
	\setlength{\labelwidth}{8mm}
	\setlength{\itemsep}{0mm}}}
	{\end{list}}

@@

\newcommand{\yes}{\text{yes}}
\newcommand{\no}{\text{no}}
\newcommand{\opt}{\text{opt}}
\newcommand{\LOCC}{\text{LOCC}}

@@

\DeclareFixedFont{\ttb}{T1}{txtt}{bx}{n}{10}
\DeclareFixedFont{\ttm}{T1}{txtt}{m}{n}{10}
\definecolor{deepblue}{rgb}{0,0,0.5}
\definecolor{deepred}{rgb}{0.6,0,0}
\definecolor{deepgreen}{rgb}{0,0.5,0}
\newcommand\cppstyle{\lstset{
language=C++,
basicstyle=\ttm,
otherkeywords={uint8_t, __m256i, size_t, ASSERT_TRUE, EXPECT_TRUE, TEST, BENCHMARK},
keywordstyle=\ttb\color{deepblue},
emphstyle=\ttb\color{deepblue},
stringstyle=\color{deepgreen},
commentstyle=\fontfamily{txtt}\selectfont\color{gray},
showstringspaces=false,
literate={*}{{\char42}}1
         {-}{{\char45}}1
}}
\lstnewenvironment{cpp}[1][]
{\cppstyle\lstset{#1}}{}
\newcommand\pythonstyle{\lstset{
language=python,
basicstyle=\ttm,
morekeywords={assert,as,echo},
keywordstyle=\ttb\color{deepblue},
emphstyle=\ttb\color{deepblue},
stringstyle=\color{deepgreen},
commentstyle=\fontfamily{txtt}\selectfont\color{gray},
showstringspaces=false,
literate={*}{{\char42}}1
         {-}{{\char45}}1
}}
\lstnewenvironment{python}[1][]
{\pythonstyle\lstset{#1}}{}

\makeatletter
\let\@fnsymbol\@arabic
\makeatother

%-----------------------------------------------------------------------------%
% Main document
%-----------------------------------------------------------------------------%

\begin{document}

\title{\LARGE\bf \qustop: Quantum state optimizer}

\author[1]{Vincent Russo}

\renewcommand\Affilfont{\normalsize\itshape}
\renewcommand\Authfont{\large}
\setlength{\affilsep}{6mm}
\renewcommand\Authand{\rule{10mm}{0mm}}

@@

\begin{abstract}
    The \qustop\ package is an open-source Python module for studying various
    quantum state optimization scenarios including calculating optimal values
    for quantum state distinguishability, quantum state exclusion, and quantum
    state cloning.
\end{abstract}

%------------------------------------------------------------------------------%
\section{Introduction}
\label{sec:intro}
%------------------------------------------------------------------------------%

%------------------------------------------------------------------------------%
\section{Definitions}
\label{sec:definitions}
%------------------------------------------------------------------------------%

It holds that
\begin{equation}
    0 \leq \omega_t(\eta, \LOCC) \leq \omega_t(\eta, \SEP) \leq \omega_t(\eta, \PPT) \leq \omega_t(\eta, \POS) \leq 1
\end{equation}
for all $t \in [U, M]$.

It holds that
\begin{equation}
    0 \leq \omega_U(\eta, T) \leq \omega_M(\eta, T) \leq 1
\end{equation}
for all $T \in [\LOCC, \SEP, \PPT, \POS]$.

The four Bell states $\ket{\psi_0}, \ldots, \ket{\psi_3} \in \complex^2 \otimes
\complex^2$ are defined as

\begin{equation}
    \begin{aligned}
        \ket{\psi_0} = \frac{1}{\sqrt{2}} \left(\ket{00} + \ket{11} \right), &\quad
        \ket{\psi_1} = \frac{1}{\sqrt{2}} \left(\ket{01} + \ket{10} \right), \\
        \ket{\psi_2} = \frac{1}{\sqrt{2}} \left(\ket{01} - \ket{10} \right), &\quad
        \ket{\psi_3} = \frac{1}{\sqrt{2}} \left(\ket{00} - \ket{11} \right).
    \end{aligned}
\end{equation}

\subsection{Minimum-error distinguishability of four two-qubit states via PPT measurements}

\begin{theorem}\label{thm:min_error_four_state_two_qubit_ppt}
    Let $\X = \Y = \complex^2$. For any PPT measurement $\{ P_1, P_2, P_3, P_4
    \} \subset \PPT(\X : \Y)$, the success probability of correctly
    discriminating the states corresponding to the set 
	\begin{equation}
        \left\{ 
            \ket{\phi_n^{(1)}},
            \ket{\phi_n^{(2)}},
            \ket{\phi_m^{(3)}},
            \ket{\phi_m^{(4)}} 
        \right\} \subset 
        \left(\X \otimes \Y \right)
	\end{equation}
    assuming a uniform distribution $p_1 = p_2 = p_3 = p_4 = 1/4$, is at most
    \begin{equation}
        \frac{1}{2} \left(1 + \frac{n+m}{2}\right)
    \end{equation}
    for all $n \in [0, 1]$ and $m \in [0, 1]$.
\end{theorem}
\begin{proof}
	Define the Hermitian operator
	\begin{equation}
        H_{n,m} = \frac{1}{2}
        \begin{pmatrix}
            1+n & 0 & 0 & 0 \\
            0 & 1+m & 0 & 0 \\
            0 & 0 & 1+m & 0 \\
            0 & 0 & 0 & 1+n
        \end{pmatrix}
        \in \Herm(\X \otimes \Y).
	\end{equation}
	It holds that
	\begin{equation}
        \frac{1}{4}\tr(H_{n,m}) = \frac{1}{2} \left(1 + \frac{n+m}{2}\right).
	\end{equation}
    We now desire to show that 
    \begin{equation}
        \begin{aligned}
            H_{n,m} - \pt_{\X}\left(\phi_n^{(1)}\right) \in \Pos(\X \otimes\Y),
            &\quad
            H_{n,m} - \pt_{\X}\left(\phi_n^{(2)}\right) \in \Pos(\X \otimes\Y), \\
            H_{n,m} - \pt_{\X}\left(\phi_m^{(3)}\right) \in \Pos(\X \otimes\Y),
            &\quad
            H_{n,m} - \pt_{\X}\left(\phi_n^{(4)}\right) \in \Pos(\X \otimes\Y),
        \end{aligned}
    \end{equation}
    for all $n \in [0, 1]$ and $m \in [0, 1]$. Observe that
    \begin{equation}\label{eq:partial-transpose-states}
        \begin{aligned}
            \pt_{\X}\left(\phi_n^{(1)}\right) = 
            \begin{pmatrix}
                \alpha^2 & 0 & 0 & 0 \\
                0 & 0 & \alpha\beta & 0 \\
                0 & \alpha\beta & 0 & 0 \\
                0 & 0 & 0 & \beta^2
            \end{pmatrix}, &\quad 
            \pt_{\X}\left(\phi_n^{(2)}\right) = 
            \begin{pmatrix}
                \beta^2 & 0 & 0 & 0 \\
                0 & 0 & -\alpha\beta & 0 \\
                0 & -\alpha\beta & 0 & 0 \\
                0 & 0 & 0 & \alpha^2
            \end{pmatrix}, \\ 
            \pt_{\X}\left(\phi_m^{(3)}\right) = 
            \begin{pmatrix}
                0 & 0 & 0 & \gamma\delta \\
                0 & \gamma^2 & 0 & 0 \\
                0 & 0 & \delta^2 & 0 \\
                \gamma\delta & 0 & 0 & 0
            \end{pmatrix}, &\quad 
            \pt_{\X} \left(\phi_m^{(4)}\right) = 
            \begin{pmatrix}
                0 & 0 & 0 & -\gamma\delta \\
                0 & \delta^2 & 0 & 0 \\
                0 & 0 & \gamma^2 & 0 \\
                -\gamma\delta & 0 & 0 & 0
            \end{pmatrix}.
        \end{aligned}
    \end{equation}
    We can represent $H_{n,m}$ as
    \begin{equation}
        H_{n,m} = 
        \begin{pmatrix}
            \alpha^2 & 0 & 0 & 0 \\
            0 & \gamma^2 & 0 & 0 \\
            0 & 0 & \gamma^2 & 0 \\
            0 & 0 & 0 & \alpha^2
        \end{pmatrix}
    \end{equation}
    since it holds that
    \begin{equation}
        \alpha^2 = \left(\sqrt{\frac{1+n}{2}}\right)^2 = \frac{1 + n}{2} 
        \quad \text{and} \quad 
        \gamma^2 = \left(\sqrt{\frac{1+m}{2}}\right)^2 = \frac{1 + m}{2}.
    \end{equation}
    Consider $k=1$ and observe that
    \begin{equation}
        H_{n,m} - \pt_{\X}\left(\phi_n^{(1)}\right) = 
        \begin{pmatrix}
            \alpha^2 - \alpha^2 & 0 & 0 & 0 \\
            0 & \gamma^2 & -\alpha\beta & 0 \\
            0 & -\alpha\beta & \gamma^2 & 0 \\
            0 & 0 & 0 & \alpha^2 - \beta^2
        \end{pmatrix}
    \end{equation}
    which is positive semidefinite as $0 \leq \beta \leq \alpha \leq 1$ and $0
    \leq \delta \leq \gamma \leq 1$ for all $n \in [0,1]$ and $m \in [0,1]$. A
    similar observation can be made for $k = 2, 3, 4$
\end{proof}

For completeness, we provide the following code listing that makes use of
\qustop to encode the setting from
Theorem~\ref{thm:min_error_four_state_two_qubit_ppt}.

\begin{figure}[!htpb]
    \centering
    \lstinputlisting[language=Python]{code/four_state_two_qubit_ppt.py}
    \caption{}
    \label{fig:ppt_four_state_two_qubit}
\end{figure}

\subsubsection*{Example: Indistinguishable maximally entangled states}

Consider the following ensemble in $\complex^4 \otimes \complex^4$ defined as
\begin{equation}\label{eq:ydy_ensemble}
    \eta = \left\{ \left(\frac{1}{4}, \ket{\psi_0} \otimes \ket{\psi_0} \right), 
                  \left(\frac{1}{4}, \ket{\psi_2} \otimes \ket{\psi_1} \right),
                  \left(\frac{1}{4}, \ket{\psi_3} \otimes \ket{\psi_1} \right),
                  \left(\frac{1}{4}, \ket{\psi_1} \otimes \ket{\psi_1} \right) 
            \right\} \subset \complex^4 \otimes \complex^4.
\end{equation}
In~\cite{yu2012four}, these states were shown to not be perfectly
distinguishable via PPT measurements, although the optimal probability with
which one is able to distinguish via PPT measurements was left as an open
question. This question was resolved in~\cite{cosentino2013positive}, where it
was shown that $\omega_M(\eta, \PPT) = 7/8$.

In~\cite{cosentino2013positive} it was also shown that $\omega_U(\eta, \PPT) = 3/4$.

\begin{figure}[!htpb]
    \centering
    \lstinputlisting[language=Python]{code/indstinguishable_mes.py}
    \caption{Calculate the minimum-error and unambiguous probabilities of
    distinguishing the ensemble from Equation~\eqref{eq:ydy_ensemble} via PPT
    measurements and separable measurements.}
    \label{fig:ppt_ydy}
\end{figure}

\subsubsection*{Example: Entanglement cost of distinguishing Bell states}

One may ask whether the ability to distinguish a state can be improved by
making use of an auxiliary resource state
\begin{equation}
    \ket{\tau_{\epsilon}} = \sqrt{\frac{1+\epsilon}{2}}\ket{00} + \sqrt{\frac{1-\epsilon}{2}},
\end{equation}
for some $\epsilon \in [0, 1]$.

\bibliographystyle{alpha}
\bibliography{refs}

\end{document}
